---
title: "figure5"
author: "Vanessa Dumeaux"
date: "12/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(VISION)
library(here)
library(ggplot2)
library("LaCroixColoR")
library(reshape2)
library(viridis)
library(viridisLite)
library(RColorBrewer)
library(here)
library(pheatmap)
library(dplyr)
library(readxl)
library(slingshot)



source(here("src", "bresat.R"))
```

## Import scvi, magic counts and VISION results


```{r}
vis.pth <- here("data", "data04-VISION")
scvi.pth <- here("data", "data03-scvi")
    
finalall <- readRDS(file.path(vis.pth, "fcz_3_6final-adjLogcounts_vis.rds"))

counts <- read.csv(file.path(scvi.pth, "fcz_3_6final-adjLogcounts_magic.csv"))
counts <- subset(counts, select = -c(X))
dim(counts)

all.latent <- finalall@LatentSpace
colnames(all.latent) <- c("UMAP1", "UMAP2")

all.metadata <- finalall@metaData
table(all.metadata$leiden_scVI_2)
table(all.metadata$leiden_scVI_2, all.metadata$drug_day)
all.metadata$leiden_scVI_2 <- recode_factor(all.metadata$leiden_scVI_2, 
                                            "0" = "cluster1", "1" = "cluster2", "2" = "cluster3",
                                            "3" = "cluster4", "others" = "others")


sig.scores <- finalall@SigScores


```



```{r}

full.df <- cbind(all.latent, all.metadata, sig.scores)


```

## Figure 5A
```{r}
library(MetBrewer)
p1 <- ggplot(subset(full.df, !leiden_scVI_2 == "others"), aes(x=UMAP1, y=UMAP2, col = leiden_scVI_2)) +
  geom_point(size = 0.4) + 
  scale_color_manual(values=met.brewer("Cross", 6)) +
  theme_void()
  
p1

ggsave("output/figure5a-all_umap_leiden2.pdf", p1, width = 5, height = 4)
```
```{r}
#Plot 1

sel <- which(all.metadata$leiden_scVI_2 %in% c("cluster1", "cluster2", "cluster3", "cluster4"))



colour_map = brewer.pal(9,'Set1')

par(xpd=TRUE)
par(mar=c(4.5,5.5,2,7))
plot(all.latent[sel, 1], all.latent[sel,2], col=colour_map[all.metadata$leiden_scVI_2[sel]], bty='L', xlab='UMAP1', ylab='UMAP2')
legend(x=15, y=7, legend=unique(all.metadata$leiden_scVI_2[sel]), fill=colour_map[as.integer(unique(all.metadata$leiden_scVI_2[sel]))])


adata_start <- slingshot(all.latent[sel,], clusterLabels = all.metadata$leiden_scVI_2[sel], start.clus="cluster2")

pdf("output/figure5a-slingshot_scvi.pdf", useDingbats=FALSE)
par(xpd=TRUE)
plot(all.latent[sel, 1], all.latent[sel,2], col=colour_map[as.factor(all.metadata$leiden_scVI_2[sel])], pch=16, asp = 1, bty='L', xlab='UMAP1', ylab='UMAP2')
lines(SlingshotDataSet(adata_start), lwd=2, type='lineages')
legend(x=10, y=20, legend=unique(all.metadata$leiden_scVI_2[sel]), fill=brewer.pal(9,'Set1')[as.integer(unique(all.metadata$leiden_scVI[sel]))])
dev.off()


```

## Figure 5D

```{r}
p2 <- ggplot(subset(full.df, !leiden_scVI_2 == "others"), aes(x=UMAP1, y=UMAP2, col = rESR_RP)) +
  geom_point(size = 0.4) + 
  scale_color_viridis_c() +
  theme_void()

ggsave("output/figure5d-all_umap_RP.pdf", p2, width = 4.5, height = 4)
```


## Figure 5E

```{r}

p3 <- ggplot(subset(full.df, !leiden_scVI_2 == "others"), aes(x=UMAP1, y=UMAP2, col = hsp2)) +
  geom_point(size = 0.4) + 
  scale_color_viridis_c() +
  theme_void()

ggsave("output/figure5e-all_umap_hsp2.pdf", p3, width = 4.5, height = 4)
```




## Identify most variable VISION signatures and explore gene expression

### Compute median signature score for each cluster and identify the most variable signatures across clusters


```{r}
library(viridis)
library(viridisLite)
library(RColorBrewer)

library(dplyr)

library(dplyr)

clusters <- all.metadata

sigtb <- data.frame(sig.scores)
sigtb$leiden_scvi <- clusters$leiden_scVI_2

sig.group <- subset(sigtb, !leiden_scvi=="others") %>%
  group_by(leiden_scvi) %>%
  summarise_at(.vars = names(.)[1:43],.funs = c(median="median"))

df.median.sig <- t(sig.group[, -1])
colnames(df.median.sig) <- sig.group$leiden_scvi


bs <- sig.ranksum(df.median.sig, ns=1:nrow(df.median.sig), full.return = TRUE)

bs.var <- apply(bs$dat, 1, var)

sel <- rownames(bs$dat)[bs.var > 0.05]
sel
```
### Check consistency of expression for most variable genes in each signature across UT cells

#### Plot correlation matrix

```{r}

library(MetBrewer)

breaksList = seq(-1, 1, by = 0.1)
sel.sig <- gsub("_median", "", sel)

pdf("output/siggenes_cormatrix_fcz3-6.pdf")

plot.cc <- lapply(sel.sig, function(x){
  
  corm <- NULL
  
  magic.cellcycle <- t(counts[, colnames(counts) %in% names(finalall@sigData[[x]]@sigDict), drop = FALSE])
  
  if(nrow(magic.cellcycle) > 0){
    
    colnames(magic.cellcycle) <- paste0("cell", 1:5865)
    sel.var <- which(apply(magic.cellcycle,1, var)>0.03)
    
    magic.cellcycle <- magic.cellcycle[sel.var, , drop = FALSE]
    
    if(nrow(magic.cellcycle) > 1){
      
      corm <- cor(t(magic.cellcycle))
      
      pheatmap(corm,
             color = rev(met.brewer("Cassatt1",n=21,type="continuous")), 
             cluster_rows = TRUE, cluster_cols = TRUE, main = x, breaks = breaksList,
             scale = "none")
    }
    }
})
dev.off()

```
### Check expression of the most variable genes in each signature
```{r}
cols <- as.character(lacroix_palette("PassionFruit", type = "continuous", n=6))
names(cols) <- levels(all.metadata$leiden_scVI_2)

pdf("output/siggenes_heatmap_fcz3-6.pdf")
plot.cc <- lapply(sel.sig, function(x){
      bs <- list()
      
      magic.cellcycle <- t(counts[, colnames(counts) %in% names(finalall@sigData[[x]]@sigDict), drop = FALSE])
      
      if(nrow(magic.cellcycle) > 0){
        colnames(magic.cellcycle) <- paste0("cell", 1:5865)
        sel.var <- which(apply(magic.cellcycle, 1, var)>0.03)
        
        magic.cellcycle <- magic.cellcycle[sel.var, , drop = FALSE]
        rownames(all.metadata) <- colnames(magic.cellcycle)
    
        if(nrow(magic.cellcycle) > 1){
          bs <- sig.ranksum(magic.cellcycle, ns=1:nrow(magic.cellcycle), full.return = TRUE)
          
          pheatmap(bs$dat, 
             annotation_col = all.metadata[bs$pat.order, colnames(all.metadata)=="leiden_scVI_2",  drop = FALSE], 
             annotation_colors = list(leiden_scVI_2=cols), 
             color = viridisLite::viridis(n=100, option="G", direction = -1), 
             cluster_rows = FALSE, cluster_cols = FALSE, main = x,
             show_colnames = FALSE, scale = "row")
          }
      }
      
      return(bs)
      })
dev.off()

names(plot.cc) <- sel.sig

names(plot.cc)[unlist(lapply(plot.cc, function(x){length(x)==0}))] # check signatures with <= 1 most variable genes
```
# Select final signatures and make figures ----------------

We selected the most variable signatures across clusters which also avoid redundancy and exhibit robustness in the the directionnality of expression for the most variable genes in signature.
```{r}

sel.paper.median <- c("hsp2_median", "alcohol_dehydrogenase_median", "rESR_RP_median")

```


## Figure 5G

```{r}

order_clusters <- c("cluster2","cluster1","cluster3", "cluster4")

df.plot.sig.heatmap <- df.median.sig[match(sel.paper.median, rownames(df.median.sig)), match(order_clusters, colnames(df.median.sig))]

      
pheatmap::pheatmap(df.plot.sig.heatmap,
             color = viridisLite::viridis(n=100, option="G", direction = -1), 
             cluster_rows = FALSE, cluster_cols = FALSE, 
             labels_row = c("heat-shock stress", 
                            "alcohol dehydrogenase",
                           "RP"),
             scale = "row", filename = "output/figure5g-heatmap_visionSig_sel_bresat.pdf", height = 3, width = 4)

```

## Figure 5H

```{r}
p4 <- ggplot(subset(full.df, !leiden_scVI_2 == "others"), aes(x=UMAP1, y=UMAP2, col = alcohol_dehydrogenase)) +
  geom_point(size = 0.4) + 
  scale_color_viridis_c() +
  theme_void()

ggsave("output/figure5h-all_umap_alcohol-dehydrogenase.pdf", p4, width = 5.5, height = 4)
```

