---
title: "figure3"
author: "Vanessa Dumeaux"
date: "12/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(VISION)
library(here)
library(ggplot2)
library("LaCroixColoR")
library(reshape2)
library(viridis)
library(viridisLite)
library(RColorBrewer)
library(here)
library(pheatmap)
library(dplyr)
library(readxl)
library(slingshot)



source(here("src", "bresat.R"))
```

## Import scvi, magic counts and VISION results


```{r}
vis.pth <- here("data", "data04-VISION")
scvi.pth <- here("data", "data03-scvi")
    
finalall <- readRDS(file.path(vis.pth, "fcz_3_6final-adjLogcounts_vis.rds"))

counts <- read.csv(file.path(scvi.pth, "fcz_3_6final-adjLogcounts_magic.csv"))
counts <- subset(counts, select = -c(X))
dim(counts)

all.latent <- finalall@LatentSpace
colnames(all.latent) <- c("UMAP1", "UMAP2")

all.metadata <- finalall@metaData
table(all.metadata$leiden_scVI_2)
table(all.metadata$leiden_scVI_2, all.metadata$drug_day)
all.metadata$leiden_scVI_2 <- recode_factor(all.metadata$leiden_scVI_2, 
                                            "0" = "cluster1", "1" = "cluster2", "2" = "cluster3",
                                            "3" = "cluster4", "others" = "others")


sig.scores <- finalall@SigScores


```



```{r}

full.df <- cbind(all.latent, all.metadata, sig.scores)


```

## Figure 4A
```{r}
library(MetBrewer)
p1 <- ggplot(subset(full.df, !leiden_scVI_2 == "others"), aes(x=UMAP1, y=UMAP2, col = leiden_scVI_2)) +
  geom_point(size = 0.4) + 
  scale_color_manual(values=met.brewer("Cross", 6)) +
  theme_void()
  
p1

ggsave("output/figure4a-all_umap_leiden2.pdf", p1, width = 5, height = 4)
```
```{r}
#Plot 1

sel <- which(all.metadata$leiden_scVI_2 %in% c("cluster1", "cluster2", "cluster3", "cluster4"))



colour_map = brewer.pal(9,'Set1')

par(xpd=TRUE)
par(mar=c(4.5,5.5,2,7))
plot(all.latent[sel, 1], all.latent[sel,2], col=colour_map[all.metadata$leiden_scVI_2[sel]], bty='L', xlab='UMAP1', ylab='UMAP2')
legend(x=15, y=7, legend=unique(all.metadata$leiden_scVI_2[sel]), fill=colour_map[as.integer(unique(all.metadata$leiden_scVI_2[sel]))])


adata_start <- slingshot(all.latent[sel,], clusterLabels = all.metadata$leiden_scVI_2[sel], start.clus="cluster2")

pdf("output/figure4a-slingshot_scvi.pdf", useDingbats=FALSE)
par(xpd=TRUE)
plot(all.latent[sel, 1], all.latent[sel,2], col=colour_map[as.factor(all.metadata$leiden_scVI_2[sel])], pch=16, asp = 1, bty='L', xlab='UMAP1', ylab='UMAP2')
lines(SlingshotDataSet(adata_start), lwd=2, type='lineages')
legend(x=10, y=20, legend=unique(all.metadata$leiden_scVI_2[sel]), fill=brewer.pal(9,'Set1')[as.integer(unique(all.metadata$leiden_scVI[sel]))])
dev.off()


```

## Figure 4C

```{r}
p2 <- ggplot(subset(full.df, !leiden_scVI_2 == "others"), aes(x=UMAP1, y=UMAP2, col = rESR_RP)) +
  geom_point(size = 0.4) + 
  scale_color_viridis_c() +
  theme_void()

ggsave("output/figure4c-all_umap_RP.pdf", p2, width = 4.5, height = 4)
```


## Figure 4D

```{r}

p3 <- ggplot(subset(full.df, !leiden_scVI_2 == "others"), aes(x=UMAP1, y=UMAP2, col = hsp2)) +
  geom_point(size = 0.4) + 
  scale_color_viridis_c() +
  theme_void()

ggsave("output/figure4d-all_umap_hsp2.pdf", p3, width = 4.5, height = 4)
```



# Figure 4E

```{r}
library("LaCroixColoR")
library(dplyr)

df.dotplot <- full.df[, c("drug_day", "leiden_scVI_2")]


df.dotplot <- subset(df.dotplot, !leiden_scVI_2 == "others") %>%
  group_by(drug_day, leiden_scVI_2) %>%
  summarise (ct = n()) %>%
  group_by(drug_day) %>%
  mutate(freq = ct / sum(ct)) %>%
  group_by(leiden_scVI_2) %>%
  mutate(freqcluster = ct / sum(ct))



df.dotplot$drug_day <- factor(df.dotplot$drug_day, 
                             levels = c("FCZ_3", "FCZ_6"), ordered = TRUE)
df.dotplot$leiden_scVI_2 <- factor(df.dotplot$leiden_scVI_2, levels = rev(levels(df.dotplot$leiden_scVI_2)), ordered = TRUE)


p4 <- ggplot(df.dotplot, aes(y=drug_day, x = leiden_scVI_2,
                         size=freqcluster, colour = freq)) +
  geom_point() +
  coord_flip() +
  theme_minimal() +
  scale_color_viridis_c(option = "inferno") +
  theme(legend.position="top", legend.box = "vertical", legend.justification='left',
        legend.key.height= unit(0.2, 'cm'),
        legend.key.width= unit(0.2, 'cm'),
        panel.grid.major = element_line(size=0.5),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size=6),
        text = element_text(size=3),
        axis.title = element_blank(),
        title =element_text(size=4))
ggsave("output/figure4e-dotplot_clusters.pdf", p4, width = 3, height = 3)

```

## Identify most variable VISION signatures and explore gene expression

### Compute median signature score for each cluster and identify the most variable signatures across clusters


```{r}
library(viridis)
library(viridisLite)
library(RColorBrewer)

library(dplyr)

library(dplyr)

clusters <- all.metadata

sigtb <- data.frame(sig.scores)
sigtb$leiden_scvi <- clusters$leiden_scVI_2

sig.group <- subset(sigtb, !leiden_scvi=="others") %>%
  group_by(leiden_scvi) %>%
  summarise_at(.vars = names(.)[1:43],.funs = c(median="median"))

df.median.sig <- t(sig.group[, -1])
colnames(df.median.sig) <- sig.group$leiden_scvi


bs <- sig.ranksum(df.median.sig, ns=1:nrow(df.median.sig), full.return = TRUE)

bs.var <- apply(bs$dat, 1, var)

sel <- rownames(bs$dat)[bs.var > 0.05]
sel
```
### Check consistency of expression for most variable genes in each signature across UT cells

#### Plot correlation matrix

```{r}

library(MetBrewer)

breaksList = seq(-1, 1, by = 0.1)
sel.sig <- gsub("_median", "", sel)

pdf("output/siggenes_cormatrix_fcz3-6.pdf")

plot.cc <- lapply(sel.sig, function(x){
  
  corm <- NULL
  
  magic.cellcycle <- t(counts[, colnames(counts) %in% names(finalall@sigData[[x]]@sigDict), drop = FALSE])
  
  if(nrow(magic.cellcycle) > 0){
    
    colnames(magic.cellcycle) <- paste0("cell", 1:5865)
    sel.var <- which(apply(magic.cellcycle,1, var)>0.03)
    
    magic.cellcycle <- magic.cellcycle[sel.var, , drop = FALSE]
    
    if(nrow(magic.cellcycle) > 1){
      
      corm <- cor(t(magic.cellcycle))
      
      pheatmap(corm,
             color = rev(met.brewer("Cassatt1",n=21,type="continuous")), 
             cluster_rows = TRUE, cluster_cols = TRUE, main = x, breaks = breaksList,
             scale = "none")
    }
    }
})
dev.off()

```
### Check expression of the most variable genes in each signature
```{r}
cols <- as.character(lacroix_palette("PassionFruit", type = "continuous", n=6))
names(cols) <- levels(all.metadata$leiden_scVI_2)

pdf("output/siggenes_heatmap_fcz3-6.pdf")
plot.cc <- lapply(sel.sig, function(x){
      bs <- list()
      
      magic.cellcycle <- t(counts[, colnames(counts) %in% names(finalall@sigData[[x]]@sigDict), drop = FALSE])
      
      if(nrow(magic.cellcycle) > 0){
        colnames(magic.cellcycle) <- paste0("cell", 1:5865)
        sel.var <- which(apply(magic.cellcycle, 1, var)>0.03)
        
        magic.cellcycle <- magic.cellcycle[sel.var, , drop = FALSE]
        rownames(all.metadata) <- colnames(magic.cellcycle)
    
        if(nrow(magic.cellcycle) > 1){
          bs <- sig.ranksum(magic.cellcycle, ns=1:nrow(magic.cellcycle), full.return = TRUE)
          
          pheatmap(bs$dat, 
             annotation_col = all.metadata[bs$pat.order, colnames(all.metadata)=="leiden_scVI_2",  drop = FALSE], 
             annotation_colors = list(leiden_scVI_2=cols), 
             color = viridisLite::viridis(n=100, option="G", direction = -1), 
             cluster_rows = FALSE, cluster_cols = FALSE, main = x,
             show_colnames = FALSE, scale = "row")
          }
      }
      
      return(bs)
      })
dev.off()

names(plot.cc) <- sel.sig

names(plot.cc)[unlist(lapply(plot.cc, function(x){length(x)==0}))] # check signatures with <= 1 most variable genes
```
# Select final signatures and make figures ----------------

We selected the most variable signatures across clusters which also avoid redundancy and exhibit robustness in the the directionnality of expression for the most variable genes in signature.
```{r}

sel.paper.median <- c("hsp2_median", "alcohol_dehydrogenase_median", "rESR_RP_median")

```


## Figure 4G

```{r}

order_clusters <- c("cluster2","cluster1","cluster3", "cluster4")

df.plot.sig.heatmap <- df.median.sig[match(sel.paper.median, rownames(df.median.sig)), match(order_clusters, colnames(df.median.sig))]

      
pheatmap::pheatmap(df.plot.sig.heatmap,
             color = viridisLite::viridis(n=100, option="G", direction = -1), 
             cluster_rows = FALSE, cluster_cols = FALSE, 
             labels_row = c("heat-shock stress", 
                            "alcohol dehydrogenase",
                           "RP"),
             scale = "row", filename = "output/figure4g-heatmap_visionSig_sel_bresat.pdf", height = 3, width = 4)

```

## Figure 4H

```{r}
p4 <- ggplot(subset(full.df, !leiden_scVI_2 == "others"), aes(x=UMAP1, y=UMAP2, col = alcohol_dehydrogenase)) +
  geom_point(size = 0.4) + 
  scale_color_viridis_c() +
  theme_void()

ggsave("output/figure4h-all_umap_alcohol-dehydrogenase.pdf", p4, width = 5.5, height = 4)
```


# Plot growth curves ------------

## Load od data
```{r}
od <- read.csv(here("data", "od", "drug-control-od.csv"))

od$treatment <- recode_factor(od$treatment, "control" = "UT", "fcz" = "FCZ", "csp" = "CSP", "rapa" = "RAPA")
```

## Figure 3A
### plot growth curves for UT and FCZ experiments (part 1 = day 1-3 )
```{r}
## compute mean and sd OD
df.fcz <- subset(od, experiment %in% c("5", "6", "7")) %>%
  group_by(treatment, day, part, batch) %>%
  summarise(mean_od = mean(od),
            sd_od = sd(od)) %>%
  ungroup()

## add day0 timepoint
day0.df <- data.frame(treatment = c("UT", "FCZ"),
                     day = c(3,3),
                     part = c("second", "second"),
                     batch = c("second", "second"),
                     mean_od = c(0.5, 0.5),
                     sd_od = c(0,0))

df.fcz.day0 <- rbind(df.fcz, day0.df)

## make plot
mean_od_fcz <- ggplot(subset(df.fcz.day0, part == "second" & batch == "second"), aes(x=day, y=mean_od))+
  geom_line(aes(color=treatment)) + 
  geom_point(aes(color = treatment)) +
  geom_errorbar(aes(ymin=mean_od-sd_od, ymax=mean_od+sd_od, color = treatment), width=.2,
                 position=position_dodge(0.05)) +
  scale_color_manual(values=c(lacroix_palette("Coconut", type = "continuous", n=15)[15],
                     lacroix_palette("Coconut", type = "continuous", n=21)[3])) +
  labs(y="OD600") +
  theme(panel.background = element_rect(fill = "white", colour = NA), 
        panel.border = element_blank(), panel.grid = element_line(colour = "grey92"), 
        panel.grid.minor = element_blank(),
        strip.background = element_rect(fill = "grey85", colour = "grey20"), 
        legend.key = element_rect(fill = "white", colour = NA), 
        legend.title = element_blank(),
        axis.title.y = element_text(angle = 90), complete = TRUE)


mean_od_fcz

ggsave("output/figure4i-mean_od_fcz.pdf", mean_od_fcz, height = 2, width = 3.5)
```