---
title: "download raw fastq files"
author: "Vanessa Dumeaux"
date: "4/19/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Download raw fastq files -----------------

In order to download the SRA files onto your machine, we use the NCBI’s SRA toolkit, which lets us use the command line to download a specified SRA run. If you are using a Linux platform, you can type: apt install sra-toolkit in your command line to install the toolkit. You can read more about SRA toolkit here: https://www.ncbi.nlm.nih.gov/books/NBK242621/ and at their github repo: https://github.com/ncbi/sra-tools.

The toolkit works by first using the prefetch command to download the SRA file associated with the specified SRA run ID.

```{r}
system(capture.output(cat("prefetch SRRxxxx")))
```

The file SRRxxx.sra should be downloaded into your home directory at ~/ncbi/public/sra/. After you have downloaded the SRA file, you can use the command fastq-dump to extract the contents of it into a .fastq.gz file. The Edwards lab at SDSU provides a nice tutorial for how to use fastq-dump here: https://edwards.sdsu.edu/research/fastq-dump/, with each of the settings explained. A sample command to extract SRRxxxx.sra would be:

```{r}
fastq_dir <- here("data", "data01-preprocessing", "fastq_input")

if (!file.exists(fastq_dir)){
  dir.create(fastq_dir, showWarnings = FALSE) 
}

system(capture.output(cat("fastq-dump --outdir", fastq_dir,
                          "--gzip --skip-technical  --readids --read-filter pass --dumpbase --split-3 --clip ~/ncbi/public/sra/SRRxxx.sra")))
```
